\chapter*{Abstract}
%\lettrine{A}{bstract}
Machine Learning~(ML) is a fascinating field of research. In the era of
knowledge, being able to find the right information in enormous amounts of data
(e.g., the internet) and summarize it in a form that is compact and yet retains
all the content one is interested in, is a key factor of success or failure in
many fields. I am particularly interested in applying ML to vision problems
because we, as humans, rely heavily on vision for our daily operations.
Improvements in the technology at our disposal to interpret visual data can
have a direct and remarkably rapid impact on many practical applications such
as assist or automate driving, analyze medical images, aid surgeons in the
operating room or improve the quality of life for visually impaired people.
This manuscript presents my work on Recurrent Neural Networks~(RNNs) and
RNN-based models applied to visual data, describing three models I proposed,
namely ReNet, ReSeg and DEConvLSTM. The first is an RNN-based alternative to
Convolutional Neural Networks~(CNNs) for object classification. The carefully
designed interaction between the RNNs in the architecture allows the model to
capture the full context of the image in just two levels of hierarchy as
opposed to the many layers typically required by CNN-based models. The
evolution of this model for semantic segmentation, called ReSeg, takes
advantage of a similar inner structure as ReNet, further improved by the
adoption of pretrained CNNs as well as the addition of transposed convolutional
layers. Finally, the DEConvLSTM architecture addresses the much harder task of
semantic segmentation in videos.  To address this task I proposed a model that
merges direct convolutions, transposed convolutions and RNNs in a unique
coherent structure. The DEConvLSTM model exploits the speed of CNNs to process
spatial information and the ability of RNNs to retain information through
several steps of computation, and proved to be a valid architecture for video
semantic segmentation. For each model, the architecture is first presented in
detail, followed by a description of the experimental settings and of the
datasets used for its evaluation. Results on publicly available dataset are
compared to the state-of-the-art and discussed thoroughly.
